---
title: "Herrschaft_Ryan_Assignment4_R_Markdown"
output: html_notebook
---
Ryan Herrschaft, 11/21/19
This is the R-script for assignment 4
 
First I start out with some code to set up the document.
```{r}
#Clean out environment
rm(list = ls())


#Session info
date()
sessionInfo()

#install needed packages
#install.packages("dslabs")
#install.packages ("tidyverse")
#install.packages("Hmisc")
library(Hmisc)
library (dslabs)
library (tidyverse)

bednet <- read_csv("C:/Users/rherrsch/Downloads/ghana_bednet.csv")
bednet
```
Questions 1 & 2: There are eight variables and 26,342 observations
```{r}
describe(bednet)
#Questions 1 & 2: There are eight variables and 26,342 observations

summary(bednet)
```


3.)What proportion of the children represented in the dataset are male?

```{r}
#Note, 0 is male, 1 is female
#Need to clean sex variable

bednet$sex[bednet$sex=="F"] <- 1

bednet$sex[bednet$sex=="."] <- NA

describe(bednet$sex)
```
13305/26,341 or 55.05% of children are male.




4.) What is the total number of years of follow-up experienced by female children in intervention areas?
```{r}


#Note that the variable "bednet" says if the child was in an intervention area. 0=control, 1=intervention
#Also note that for sex variable 1=female
#and, follyr is the continuous follow-up variable

female_only <- filter(bednet, sex == 1, bednet == 1)
describe (female_only$follyr)

#There was a missing value that had to be accounted for.

sum(female_only$follyr, na.rm = TRUE)


```
The total number of years is 8395.184

5.)What is the proportion of all children in cluster 52 who died during followup?
```{r}
# Note that outcome is the variable that tells you if a child died 0=no, 1=yes

cluster_52_only <- filter(bednet, cluster == "52")
describe (cluster_52_only$outcome)

```
9 out of the 370 children died. This is 2.43% of children in cluster 52.


6.) What is the mean and standard deviation of the subjects age?
```{r}
mean(bednet$agemn)
sd(bednet$agemn)
```

The mean age is 25.29. The standard deviation of age is 18.025.


7.  Calculate the mortality rate (number or deaths per person year of follow up) in the control arm and, separately, in the intervention arm.
```{r}
#bednet is control/intervention varaible. 0 means control, 1 means intervention

control_only <- filter(bednet, bednet ==0)
intervention_only <- filter(bednet, bednet==1)


#mortality rate per person per year can be calculated by (number of deaths) over follyr(person years number of followup)
#because deaths are coded as 1, we can count deaths by taking the sum of the outcome variable

sum_deaths_control <-sum(control_only$outcome)
sum_deaths_control
#there were 461 deaths in the control group.

#from here, we find the number of deaths in the control sample. We take this number and divide it by the sum of follyr.

sum_follow_year_control <- sum(control_only$follyr, na.rm = TRUE)
sum_follow_year_control

#Finally, we take the number of deaths and divide by sum of follyr
mort_rate_control <-(sum_deaths_control)/(sum_follow_year_control)
mort_rate_control
#The mortality person rate per person per year for the control group is .02795


#We now repeat these steps but for the intervention group.
sum_deaths_intervention <-sum(intervention_only$outcome)
sum_deaths_intervention
#there were 396 deaths in the intervention group.

#from here, we find the number of deaths in the intervention sample. We take this number and divide it by the sum of follyr.
sum_follow_year_intervention <- sum(intervention_only$follyr, na.rm = TRUE)
sum_follow_year_intervention
#Finally, we take the number of deaths and divide by sum of follyr
mort_rate_intervention <-(sum_deaths_intervention)/(sum_follow_year_intervention)
mort_rate_intervention

# The morality rate per person per year for the intervention group is  0.02351

```
The mortality person rate per person per year for the control group is .02795

The morality rate per person per year for the intervention group is  0.02351


8. Make a bar chart showing the proportion of participants who died in each cluster.
```{r}
#I had trouble using the "group by" command to generate these results. Instead, I used aggregate

mean_death_by_cluster <- aggregate(bednet$outcome, by=list(bednet$cluster), FUN=mean) %>% rename(cluster_group = Group.1) %>% rename(Prop_death = x)

mean_death_by_cluster

barplot1<-ggplot(data= mean_death_by_cluster, aes(x= cluster_group, y= Prop_death)) + geom_bar(stat="identity") +ggtitle("Proportion of participants who died in each cluster") +labs(y= "Proportion of Deaths", x="Cluster Number") 
barplot1
```

9.  Use the glm() function from base R's stats package, and the formula 'out- come  ~ bednet'with the family argument specified as 'binomial(link = "logit")' to fit a logistic regression. Print the summary of this regression in your document.
```{r}
model <- glm(outcome ~ bednet, data=bednet, binomial(link= "logit"))
summary (model) 
print(model)
```

10. Was being in an intervention cluster (i.e. having bednets) associated with lower or higher
probability of a child dying during the course of the trial?
What was the coefficient (or odds ratio (specify which)) and p-value for this conclusion.

```{r}
exp(coef(model))
```
The odds ratio is 0.836 (p<.05), meaning that those with bednets had a decreased chance of having a child die during the course of the trial.

